\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

In the contemporary digital learning landscape, the challenge of delivering
personalized, accessible, and time-efficient education has become increasingly
critical. Modern learners face a fundamental paradox: unprecedented access to
knowledge paired with diminishing time to engage with it. This project presents
\textbf{Learnova} --- an AI-powered Audiobook and E-Learning platform designed
to bridge the gap between fragmented audiobook services and structured
educational environments through intelligent automation, audio-first delivery,
and interactive course support.

The platform enables users to transform PDF books and educational materials into
immersive, high-quality audio experiences via a state-of-the-art Text-to-Speech
(TTS) pipeline, delivering a Spotify-inspired listening interface with persistent
Mini Player support, adjustable playback speed, and chapter-level navigation.
In parallel, Learnova provides a structured course management system featuring
video lessons with AI-generated transcripts, enabling learners to search, review,
and interact with course content through natural language. The
\texttt{TranscribeAllLessons} pipeline automatically processes video lessons
through an AI transcription service, making all course content fully searchable
and accessible.

Beyond audio and course delivery, Learnova integrates a suite of six Python AI
microservices covering the full intelligent library pipeline: semantic
vectorisation using \texttt{all-mpnet-base-v2} embeddings stored in Pinecone,
a FLAN-T5-based map-reduce summarisation service, a Retrieval-Augmented
Generation (RAG) Q\&A service powered by GPT-4o-mini with Server-Sent Events
streaming and page-level citation, a hybrid recommendation engine combining
Neural Collaborative Filtering (NCF) with content-based filtering, and a
semantic search service with intent-aware re-ranking. A context-aware
conversational chatbot provides on-demand assistance grounded in book content,
ensuring each user receives accurate, citation-backed answers.

The system is built on a modern cross-platform mobile architecture using
\textbf{Flutter} and \textbf{Dart} for the frontend with BLoC state management,
backed by a \textbf{PHP Laravel 12} RESTful API and \textbf{MySQL 8.0} database.
Laravel acts as a central AI proxy gateway, forwarding requests to the
appropriate microservice and degrading gracefully on service unavailability.
An administrative panel built with FilamentPHP 3 provides full content
management, user role control, and AI metadata oversight. Secure authentication
is handled through JWT and Laravel Sanctum, with support for OTP-based password
reset and social login via Google and Facebook OAuth.

This project demonstrates the practical feasibility of unifying audio-first
consumption, structured course progression, and a full AI knowledge pipeline
within a single cohesive platform, establishing Learnova as a foundational
model for next-generation educational systems that prioritize accessibility,
personalization, and learner autonomy.

\newpage